Supplementary material for ``Screening Sinkhorn Algorithm for Regularized Optimal Transport''

### 2019-05-22 tex/v3/section_5_numeric_expes.tex
update

\begin{figure*}[t]
	\begin{center}
		% \includegraphics[width=6cm]{./figs/norm_M_Mu_marginals_toy_n1000} \hspace*{9pt}
		% \includegraphics[width=6cm]{./figs/norm_M_Nu_marginals_toy_n1000} \\
		% \includegraphics[width=6cm]{./figs/norm_M_time_toy_n1000} \hspace*{9pt}
		% \includegraphics[width=6cm]{./figs/norm_M_div_toy_n1000}
		% 
		\includegraphics[width=0.35\textwidth]{./figs/norm_M_Mu_marginals_toy_n1000}\hspace{2cm}
		\includegraphics[width=0.35\textwidth]{./figs/norm_M_Nu_marginals_toy_n1000} \\
		\includegraphics[width=0.35\textwidth]{./figs/norm_M_time_toy_n1000}\hspace{2cm}
		\includegraphics[width=0.35\textwidth]{./figs/norm_M_div_toy_n1000}
	\end{center}
	\caption{Empirical evaluation of \emph{Screenkhorn} vs Sinkhorn for normalized cost matrix. Top panel: marginal violations in relation with the budget of points. Bottom panel: (left) ratio of computation times    $\frac{T_{\text{Sinkhorn}}}{T_{\text{\emph{Screenkhorn}}}}$ and, (right) relative divergence variation. The results are averaged over $30$ trials.} 
		\label{fig:margin_expe}
\end{figure*}
%
\emph{Screenkhorn} provides good approximation of the marginals $\mu$ and $\nu$ for ``high'' values of the regularization parameter $\eta$ ($\eta > 1$). The approximation quality diminishes for small $\eta$. As expected $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ converge towards zero when increasing the budget of points. Remarkably marginal violations are almost negligible whatever the budget for high $\eta$.  According to computation gain, \emph{Screenkhorn} is, up to $2$ times, faster than Sinkhorn at low budget $n_b/n$ while the reverse holds when $n_b/n$ gets close to $1$.  Computational benefit of \emph{Screenkhorn} also depends on $\eta$ with appropriate values $\eta \leq 1$. Finally except for $\eta=0.1$ \emph{Screenkhorn} achieves a close divergence $\inr{C, P}$ compared to Sinkhorn showing that our static screening test does not harm Sinkhorn divergence.

@@

In this subsection, we analyse the impact of using \emph{Screenkhorn}
instead of Sinkhorn in a complex machine learning pipeline. Our two applications
are a dimensionality reduction technique, denoted as Wasserstein Discriminant Analysis (WDA), based on Wasserstein distance approximated
through Sinkhorn divergence \citep{flamary2018WDA} and a domain-adaptation using optimal transport mapping \citep{courty2017optimal}, named OTDA.

@@

\begin{figure*}[t]
	\centering
%	\includegraphics[width=4.cm]{./figs/wda_accur_toy.pdf}
%	\includegraphics[width=4.cm]{../figs/wda_accur_mnist.pdf}
	\includegraphics[width=0.39\textwidth]{./figs/wda_gain_toy.pdf}\hspace{1.5cm}
	\includegraphics[width=0.39\textwidth]{./figs/wda_gain_mnist.pdf}
	\caption{Wasserstein Discriminant Analysis : running time gain for a toy dataset and for mnist as a function of the number of examples and the data decimation factor in \emph{Screenkhorn}.}
	\label{fig:wda}
\end{figure*}
\begin{figure*}[t]
	\centering

	\includegraphics[width=0.43\textwidth]{./figs/da_gain_mnist_regcl1.pdf}\hspace{1cm}
	\includegraphics[width=0.43\textwidth]{./figs/da_gain_mnist_regcl10.pdf}
	\caption{OT Domain adaptation : running time gain for a toy dataset and for mnist as a function of the number of examples and the data decimation factor in \emph{Screenkhorn}.}
	\label{fig:otda}
\end{figure*}

>>>>

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=4.5cm]{./figs/norm_M_Mu_marginals_toy_n1000} \hspace*{9pt}
		\includegraphics[width=4.5cm]{./figs/norm_M_Nu_marginals_toy_n1000} \\
		\includegraphics[width=4.5cm]{./figs/norm_M_time_toy_n1000} \hspace*{9pt}
		\includegraphics[width=4.5cm]{./figs/norm_M_div_toy_n1000}
	\end{center}
	\caption{Empirical evaluation of \emph{Screenkhorn} vs Sinkhorn for normalized cost matrix. Top panel: marginal violations in relation with the budget of points. Bottom panel: (left) ratio of computation times    $\frac{T_{\text{Sinkhorn}}}{T_{\text{\emph{Screenkhorn}}}}$ and, (right) relative divergence variation. The results are averaged over 30 trials.} 
		\label{fig:margin_expe}
\end{figure*}
%
\emph{Screenkhorn} provides good approximation of the marginals $\mu$ and $\nu$ for ``high'' values of the regularization parameter $\eta$ ($\eta > 1$). The approximation quality diminishes for small $\eta$. As expected $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ converge towards zero when increasing the budget of points. Remarkably marginal violations are almost negligible whatever the budget for high $\eta$.  According to computation gain, \emph{Screenkhorn} is almost  2 times faster than Sinkhorn at high decimation factor $n/n_b$ (low budget) while the reverse holds when $n/n_b$ gets close to 1.  Computational benefit of \emph{Screenkhorn} also depends on $\eta$ with appropriate values $\eta \leq 1$. Finally except for $\eta=0.1$ \emph{Screenkhorn} achieves a  divergence $\inr{C, P}$ close to the one of Sinkhorn showing that our static screening test does not harm Sinkhorn divergence.  \emph{Screenkhorn} will be practically useful in cases when modest accuracy on the divergence is sufficient as a loss function a gradient descent method optimizes over.

@@

Here, we analyse the impact of using \emph{Screenkhorn}
instead of Sinkhorn in a complex machine learning pipeline. Our two applications
are a dimensionality reduction technique, denoted as Wasserstein Discriminant Analysis (WDA), based on Wasserstein distance approximated
through Sinkhorn divergence \citep{flamary2018WDA} and a domain-adaptation using optimal transport mapping \citep{courty2017optimal}, named OTDA.

@@

\begin{figure*}[t]
	\centering
%	\includegraphics[width=4.cm]{./figs/wda_accur_toy.pdf}
%	\includegraphics[width=4.cm]{../figs/wda_accur_mnist.pdf}
	\includegraphics[width=0.39\textwidth]{./figs/wda_gain_toy.pdf}~\hfill~
	\includegraphics[width=0.39\textwidth]{./figs/wda_gain_mnist.pdf}
	\caption{Wasserstein Discriminant Analysis : running time gain for a toy dataset and for mnist as a function of the number of examples and the data decimation factor in \emph{Screenkhorn}.}
	\label{fig:wda}
\end{figure*}
\begin{figure*}[t]
	\centering

	\includegraphics[width=0.43\textwidth]{./figs/da_gain_mnist_regcl1.pdf}~\hfill~
	\includegraphics[width=0.43\textwidth]{./figs/da_gain_mnist_regcl10.pdf}
	\caption{OT Domain adaptation : running time gain for a toy dataset and for mnist as a function of the number of examples and the data decimation factor in \emph{Screenkhorn}.}
	\label{fig:otda}
\end{figure*}
